**Loading and preprocessing the data**  
```{r ,echo=TRUE}
getwd()
setwd("C:/Users/ASEEL/Desktop")
getwd()
library(dplyr)
library(ggplot2)
library(lattice)
Data <- read.csv("activity.csv")
```

more details 
```{r ,echo=TRUE}
date <- unique(Data$date)
date
head(Data)
str(Data)
summary(Data)
summary(is.na(Data))

```
**What is mean total number of steps taken per day?**

```{r ,echo=TRUE}
par(mfrow=c(1,1))
sumD <- data.frame(date=unique(Data$date),steps=tapply(Data$steps,Data$date,sum,na.rm=TRUE))
hist(sumD$steps,xlab = "Number of Steps",main="Steps per Day",ylab="frequancy")
```
sumD is the total number of steps per eachday 
and the mean and the median for the previous data are 
```{r ,echo=TRUE}
mean(sumD$steps)
median(sumD$steps)
```
**What is the average daily activity pattern?**

```{r, echo=TRUE}
sumI <- data.frame(Interval=as.integer(unique(Data$interval)),M=tapply(Data$steps,Data$interval,mean,na.rm=TRUE))
plot(sumI$Interval,sumI$M,type="l",xlab="5-mins Intervals",ylab="Avg num of steps")
```

The 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps is 
```{r, echo=TRUE}
maxstepsI <- max(sumI$M)
maxstepsI
sumI[sumI$M==maxstepsI,]
```

** Imputing missing values**

```{r, echo=TRUE}
summary(is.na(Data$steps))
sum(is.na(Data$steps))
    ```
    
and if we make tha same  histogram without these missing data 
```{r, echo=TRUE}


sumDM <- data.frame(date=unique(Data$date),steps=tapply(Data$steps,Data$date,sum))
hist(sumDM$steps,xlab = "Number of Steps",main="Steps per Day(with missing data)",ylab="frequancy")

``` 


```{r , echo=TRUE}
Data$date <- as.Date(Data$date)
clean <- data.frame(Data[!is.na(Data$steps),])
bad <- Data[is.na(Data$steps),]
dataM <- data.frame(interval=unique(clean$interval),mean=tapply(clean$steps,clean$interval,mean))
clean2 = data.frame(merge(bad[,c("steps","date","interval")],dataM[ ,c("interval","mean")]))
clean2[,2] <- clean2[,4]
bad2<- clean2[,1:3]

newNA<- data.frame(steps=bad2[,2],date=bad2[,3],interval=bad2[,1])
newdata <- rbind(clean,newNA)
```

**Are there differences in activity patterns between weekdays and weekends?**

```{r, echo=TRUE}

Data$date <- as.Date(Data$date,"%Y %m %A")
head(Data$date)
Wday<- function(d){
  wd <- weekdays(d)
  ifelse(wd =="Saturday" | wd=="Sunday","weekend","weekday")}
date <- as.Date(date)
w <- data.frame(WD=sapply(Data$date,Wday),interval=Data$interval,steps=Data$steps)
WD <- w[w$WD== "weekday" ,]
WE <- w[w$WD== "weekend" ,]
WD <- data.frame(Interval=unique(WD$interval),steps=tapply(WD$steps,WD$interval,sum,na.rm=TRUE))
WE <- data.frame(Interval=unique(WE$interval),steps=tapply(WE$steps,WE$interval,sum,na.rm=TRUE))
par(mfrow=c(1,2))
xyplot(WD$steps~WD$Interval,type="l",xlab = "Intervals", ylab = "Steps")
xyplot(WE$steps~WE$Interval,type="l",xlab = "Intervals", ylab = "Steps")
```


